### 2017-06-14 gpu-05-03-accelerator.tex
Maxwell in (a) and (b), Pascal in (c)

\subsubsection{Analysis of GPU Parallelism}
As discussed in Section~\ref{gpu-design}, there are 10 distinct flavors of the
GPU version of the algorithm. Specifically, there is the naive strategy, the
shared strategy and 8 variations of the code generation strategy. This
section investigates the effectiveness of each flavor on NVidia's GTX~\mbox{Titan-X}
Maxwell and Pascal,
GTX980, K40c and K20m.

% TITANX
\begin{figure*}[tb]	% [htb]
  \centering
  \epsfig{file=plots/gpu-sweep-legends.eps, width=1.0\textwidth}
  \setlength\tabcolsep{0em}
  \begin{tabular}{c c c c}
  \epsfig{file=plots/titanx-maxwell-1024-strategies-w2.eps, width=0.25\textwidth} &
  \epsfig{file=plots/titanx-maxwell-2048-strategies-w2.eps, width=0.25\textwidth} &
  \epsfig{file=plots/titanx-pascal-1024-strategies-w2.eps, width=0.25\textwidth} &
  \epsfig{file=plots/k40-1024-strategies-w2.eps, width=0.25\textwidth}
  \end{tabular}
  \caption{Execution time of 1000 \textit{update\_phi} invocations across a sweep of
  	\textit{update\_phi} thread block sizes. The lower figures are a zoom-in into
	the optimal block range of the upper figures.
	Other relevant model parameters: M=4096, $|\Neighbors|$=32.
	% (a) \mbox{Titan-X(Pascal)} GPU, with kernel vectorization of width 2;
	% Relevant model parameters: K=1024, M=4096, $|\Neighbors|$=32.
	% (b) \mbox{Titan-X} GPU, without explicit kernel vectorization;
	% Relevant model parameters: K=1024, M=4096, $|\Neighbors|$=32.
  	% (c) \mbox{Titan-X} GPU, without explicit kernel vectorization;
	% K=2048, M=4096, $|\Neighbors|$=32.
	% (d) K40c GPU, with explicit kernel vectorization of width 2;
	% K=1024, M=4096, $|\Neighbors|$=32.
  }
  % \label{titanx-w1-sweep}
  \label{gpu-sweep}
\end{figure*}

@@

Figure \ref{gpu-sweep}(a) presents the performance of the Pascal \mbox{Titan-X} GPU
for all 10 strategy flavors, with an explicit vector width~2 in the kernels. The
x-axis represents different \textit{update\_phi} thread block sizes while the y-axis presents
the total execution time of 1000 invocations of the \textit{update\_phi} kernel. The
naive and shared strategies are labeled NAIVE and SHARED respectively. Further,
each flavor of the code generation strategy is labeled by GEN followed by the 3
choices that identify it.
%
The wide performance range inhibits the readability of the figure,
therefore, a zoomed-in version of Figure~\ref{gpu-sweep}(a), is provided as the bottom figure. As would
be expected, the naive strategy exhibits the worst performance over all
thread block
sizes, as it does not explicitly cache repeated device memory read operations.
The SHARED and \textit{GEN-SSS} strategies come next in terms of performance.
Both strategies cache $Grads_i$, $Probs_i$ and $\pi_i$ in shared memory but
differ in one aspect: \textit{GEN-SSS} explicitly unrolls the internal
loops of the kernel. However, there is no significant performance difference
between them. The other flavors of the code generation strategy attain higher
performance as they unroll internal loops as well as cache data in
registers. The Pascal Titan-X is fastest with the \textit{GEN-RSR} strategy and a
block size of 96 and a vector width of~2, whereas
the Maxwell Titan-X in Figure~\ref{gpu-sweep}(b) obtains the best performance with the \textit{GEN-RSS}
strategy and a thread block size of 64 and no vectorization. The results of other
vector widths
are omitted as they obtain lower performance.

@@

@@

% \looseness=-1
Figure~\ref{gpu-sweep}(c) presents the performance of the K40c GPU
for the same experimental configuration as shown for the Maxwel \mbox{Titan-X} in
Figure~\ref{gpu-sweep}(a) except that the code
is explicitly vectorized with a vector width of~2. The results
for the other code versions with vector width 4 and no vectorization are
omitted as they exhibit lower performance.
%
Surprisingly, Figure~\ref{gpu-sweep}(c) shows that the NAIVE strategy
outperforms SHARED and some of the GEN strategies. This can be explained by the
unique properties of the Tesla Super Computing line of products to which the
K40c belongs. These GPUs include enhanced L2 caching mechanisms that
accelerate repeated and sparse memory accesses. This is especially advantageous
as it caches repeated reads across streaming multiprocessors. However, the
highest performance is attained by \textit{GEN-RRS} which explicitly employs
registers for both $Probs_i$ and $Grads_i$.

The performance results obtained from the Maxwell \mbox{Titan-X} and Tesla~K40c GPUs
reinforce the importance of customizing compute kernels to each GPU's specific
architecture and capabilities. For instance, each GPU achieved its highest
performance by employing a different strategy. Moreover, each GPU displayed
different strategy-performance orderings.

>>>>

\subsubsection{Analysis of GPU Parallelism}
As discussed in Section~\ref{gpu-design}, there are 10 distinct flavors of the
GPU version of the algorithm: the naive strategy, the
shared strategy, and 8 variations of the code generation strategy. This
section investigates the effectiveness of each flavor on NVidia's GTX~\mbox{Titan-X}
(Maxwell and Pascal),
GTX980, K40c and~K20m.

% TITANX
\begin{figure*}[tb]	% [htb]
  \centering
  \epsfig{file=plots/gpu-sweep-legends.eps, width=1.0\textwidth}
  \setlength\tabcolsep{0em}
  \begin{tabular}{c c c c}
  \epsfig{file=plots/titanx-maxwell-1024-strategies-w1.eps, width=0.25\textwidth} &
  \epsfig{file=plots/titanx-maxwell-2048-strategies-w1.eps, width=0.25\textwidth} &
  \epsfig{file=plots/titanx-pascal-1024-strategies-w2.eps, width=0.25\textwidth} &
  \epsfig{file=plots/k40-1024-strategies-w2.eps, width=0.25\textwidth}
  \end{tabular}
  \caption{Execution time of 1000 \textit{update\_phi} invocations across a sweep of
  	\textit{update\_phi} thread block sizes. The lower figures are a zoom-in into
	the optimal block range of the upper figures.
	Other relevant model parameters: M=4096, $|\Neighbors|$=32.
	% (a) \mbox{Titan-X(Pascal)} GPU, with kernel vectorization of width 2;
	% Relevant model parameters: K=1024, M=4096, $|\Neighbors|$=32.
	% (b) \mbox{Titan-X} GPU, without explicit kernel vectorization;
	% Relevant model parameters: K=1024, M=4096, $|\Neighbors|$=32.
  	% (c) \mbox{Titan-X} GPU, without explicit kernel vectorization;
	% K=2048, M=4096, $|\Neighbors|$=32.
	% (d) K40c GPU, with explicit kernel vectorization of width 2;
	% K=1024, M=4096, $|\Neighbors|$=32.
  }
  % \label{titanx-w1-sweep}
  \label{gpu-sweep}
\end{figure*}

@@

Figure \ref{gpu-sweep}(a) presents the performance of the Maxwell \mbox{Titan-X} GPU
for all 10 strategy flavors, with an explicit vector width~2 in the kernels. The
x-axis represents different \textit{update\_phi} thread block sizes while the y-axis presents
the total execution time of 1000 invocations of the \textit{update\_phi} kernel. The
naive and shared strategies are labeled NAIVE and SHARED respectively. Further,
each flavor of the code generation strategy is labeled by GEN followed by the 3
choices that identify it.
%
The wide performance range inhibits the readability of the figure,
therefore, a zoomed-in version of Figure~\ref{gpu-sweep}(a) is provided as the bottom figure. As would
be expected, the naive strategy exhibits the worst performance over all
thread block
sizes, as it does not explicitly cache repeated device memory read operations.
The SHARED and \textit{GEN-SSS} strategies come next in terms of performance.
Both strategies cache $Grads_i$, $Probs_i$ and $\pi_i$ in shared memory but
differ in one aspect: \textit{GEN-SSS} explicitly unrolls the internal
loops of the kernel. However, there is no significant performance difference
between them. The other flavors of the code generation strategy attain higher
performance as they unroll internal loops as well as cache data in
registers.
The Maxwell Titan-X in Figure~\ref{gpu-sweep}(b) obtains the best performance with the \textit{GEN-RSS}
strategy and a thread block size of 64 and no vectorization. The results of other
vector widths
are omitted as they obtain lower performance.

@@

The Pascal \mbox{Titan-X} in Figure~\ref{gpu-sweep}(c) is the fastest of the GPUs
under investigation, note the y-axis range. The NAIVE strategy yields relatively
worse performance in comparison to the other GPUs. In contrast to the Maxwell
\mbox{Titan-X}, the Pascal shows very good performance with the SHARED strategy.
This GPU performs best with the \textit{GEN-RSR} strategy, a block size of~96
and a vector width of~2.

@@

% \looseness=-1
Figure~\ref{gpu-sweep}(d) presents the performance of the K40c GPU
for the same experimental configuration as shown for the Pascal \mbox{Titan-X} in
Figure~\ref{gpu-sweep}(c), again with a vector width of~2. The results
for the other code versions with vector width 4 and no vectorization are
omitted as they exhibit lower performance.
%
Surprisingly, Figure~\ref{gpu-sweep}(c) shows that the NAIVE strategy
outperforms SHARED and some of the GEN strategies. This can be explained by the
unique properties of the Tesla Super Computing line of products to which the
K40c belongs. These GPUs include enhanced L2 caching mechanisms that
accelerate repeated and sparse memory accesses. This is especially advantageous
as it caches repeated reads across streaming multiprocessors. However, the
highest performance is attained by \textit{GEN-RRS} which explicitly employs
registers for both $Probs_i$ and $Grads_i$.

These performance results for a range of GPUs
reinforce the importance of customizing compute kernels to each GPU's specific
architecture and capabilities. For instance, each GPU achieved its highest
performance by employing a different strategy. Moreover, each GPU displayed
different strategy-performance orderings.
